---
title: 'Project 3: Ovarian Cancer'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Contents
1. Overview Ovarian Cancer
2. Data Cleanup
3. Driver Mutations
4. co-existing Mutations
5. Synthetic Lethality Interactions
6. Linear Regression
7. Summary of Results
8. Conclusion

## 1. Overview Ovarian Cancer

### 1.1 General facts
Ovarian cancer is a very heterogenous disease.

## 2. Data Cleanup

first we downloaded the data and load data in R
```{r}
allDepMapData <- readRDS("~/GitHub/project-01-group-03/DepMap19Q1_allData.RDS")
```
rename the lists from the main data
```{r}
copynumber = allDepMapData[["copynumber"]]
mutation = allDepMapData[["mutation"]]
kd.ceres = allDepMapData[["kd.ceres"]]
kd.prob = allDepMapData[["kd.prob"]]
annotation = allDepMapData[["annotation"]]
expression = allDepMapData[["expression"]]

```
extract all cell lines that have ovarian cancer as primary disease. 
```{r}
annotation = annotation[which(annotation$Primary.Disease == "Ovarian Cancer"), ]
ID = annotation$DepMap_ID
expression = expression[ , which(colnames(expression) %in% ID)]
copynumber = copynumber [ , which(colnames(copynumber) %in% ID)]
kd.ceres = kd.ceres [ , which(colnames(kd.ceres) %in% ID)]
kd.prob = kd.prob [ , which(colnames(kd.prob) %in% ID)]
mutation = mutation [ ID]
```
extract mutation submatrices

remove the object allDepMapData for a clean workflow
```{r}
rm(allDepMapData)
```


## 3. Driver Mutations
juop

## 4. Co-existing Mutations

## 5. Synthetic Lethality Interactions
After identifying our most common driver mutations, we used this information to find possible synthetic lethality interaction partners

## 6. Linear Regression

## 7. Summary of Results

## 8. Conclusion

