---
title: 'Project 3: Ovarian Cancer'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Contents
1. Overview Ovarian Cancer
2. Data Cleanup
3. Driver Mutations
4. co-existing Mutations
5. Synthetic Lethality Interactions
6. Linear Regression
7. Summary of Results
8. Conclusion

## 1. Overview Ovarian Cancer

### 1.1 General facts
Ovarian cancer is a very heterogenous disease.lol

## 2. Data Cleanup

first we downloaded the data and load data in R
```{r}
allDepMapData <- readRDS("~/GitHub/project-01-group-03/DepMap19Q1_allData.RDS")
```
rename the lists from the main data
```{r}
copynumber = allDepMapData[["copynumber"]]
mutation = allDepMapData[["mutation"]]
kd.ceres = allDepMapData[["kd.ceres"]]
kd.prob = allDepMapData[["kd.prob"]]
annotation = allDepMapData[["annotation"]]
expression = allDepMapData[["expression"]]

```
extract all cell lines that have ovarian cancer as primary disease. 
```{r}
annotation = annotation[which(annotation$Primary.Disease == "Ovarian Cancer"), ]
ID = annotation$DepMap_ID
expression = expression[ , which(colnames(expression) %in% ID)]
copynumber = copynumber [ , which(colnames(copynumber) %in% ID)]
kd.ceres = kd.ceres [ , which(colnames(kd.ceres) %in% ID)]
kd.prob = kd.prob [ , which(colnames(kd.prob) %in% ID)]
mutation = mutation [ ID]
```
extract mutation submatrices

```{r}
#ich bin kurz vorm durchbruch

n = 34
#for(i in 0:n) {"ID[i]" = mutation[[i]] ;}

```


remove the object allDepMapData for a clean workflow
```{r}
rm(allDepMapData)
```

remove unnecesary cols of the annotation matrix and rename the rows with the cell lines
```{r}
rownames(annotation) = annotation$DepMap_ID
annotation = annotation[, -which(colnames(annotation) %in% c("DepMap_ID", "Aliases", "Primary.Disease", "Gender", "Source"))]

```

check the matrices if they contain any NA values and delte the containing rows

```{r}
#summary(copynumber) hier müssen wirs hinkriegen dass er den output net reinhaut


NAV = apply(copynumber, 1, function(x) {sum(is.na(x))}) 
copynumber = copynumber[-which(NAV > 0), ]  

#hier muss ich nochmal gucken ob man des machen muss weil iwi passiert da was komisches

#NAV = apply(expression, 1, function(x) {sum(is.na(x))}) 
#expression = expression[-which(NAV > 0), ]  
 
#NAV = apply(kd.ceres, 1, function(x) {sum(is.na(x))}) 
#kd.ceres = kd.ceres[-which(NAV > 0), ]

#NAV = apply(kd.prob, 1, function(x) {sum(is.na(x))}) 
#kd.prob = kd.prob[-which(NAV > 0), ]
```


## 3. Driver Mutations
juop

## 4. Co-existing Mutations

## 5. Synthetic Lethality Interactions
After identifying our most common driver mutations, we used this information to find possible synthetic lethality interaction partners

## 6. Linear Regression

## 7. Summary of Results

## 8. Conclusion

